--- 
title: "Section 1.2 TB mortality" 
author: "Mathieu Bastard, reformatted/refactored by Hazim Timimi" 
date: "`r Sys.Date()`" 
knit: (function(inputFile, encoding) {
      out_dir <- "./local";
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_dir=file.path(dirname(inputFile), out_dir))})
output:  
  html_fragment: 
    # Don’t include a table of contents 
    toc: no 
    # Set standard figure width to 12 inches 
    fig_width: 12 
    # Don’t write figure captions 
    fig_caption: FALSE 
    # Don't embed external resources (stylesheets, JS libraries) in the output 
    self_contained: FALSE 
    # Don't include <div> tags around a header and the content found until the next header
    section_divs: FALSE  


# To run this file and store output as html:
# rmarkdown::render(here::here("report/ch1-2.rmd"), output_file = "ch1-2.html", output_dir = here::here("report/local/")) 
--- 



```{r setup, include=FALSE} 
# Chapter 1
# Set options, load data, utility functions 

knitr::opts_chunk$set(echo = FALSE, 
                      results = "asis",
                      message = FALSE,
                      warning = FALSE,
                      error = TRUE, # for debugging!
                      
                      # Settings for better output of whomapper .png maps
                      fig.width = 10,
                      fig.height = 5,
                      dpi = 350
                      )

# Kill any attempt at using factors, unless we explicitly want them!
options(stringsAsFactors=FALSE)
options(scipen=999)


# Load output packages ----
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

library(data.table)
library(gtbreport)
library(here)
library(kableExtra)
library(gridExtra)
# devtools::install_github("yamanakatakuya/whomapR")
# from https://github.com/yamanakatakuya/whomapR
library(whomapR)
library(ggpubr)
library(productplots)
library(RColorBrewer)
library(tidyverse)
library(jsonlite)
library(dplyr)

# Prepare data for the chapter
source(here('report/ch1_prepare_data.R'))

# Load R functions ----
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

source(here("report/functions/html_links.R"))
source(here("report/functions/output_ggplot.R"))




# Set output options ----
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 

# Show static chart in addition to Kendo chart?
show_static= FALSE

# Save underlying data files as CSV and charts as PDF files?
pdf_csv_folder = here::here("report/local/figures/ch1-2")
save_csv = TRUE
save_pdf = TRUE
save_hidef = FALSE # TRUE only for producing a high definition map for the PDF!

# Create the output folder (only if it doesn't yet exist)
dir.create(pdf_csv_folder, showWarnings = FALSE, recursive = TRUE)

thous=function(x){
  return(paste0(x," ","000"))
}

hund <- function(x) {
  rounded_val <- round(x / 100) * 100
  formatted_val <- format(rounded_val, big.mark = " ", scientific = FALSE)
  return(formatted_val)
}


# Load standard note about Indonesia moving to WPR ----
# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
source(here("report/resources/IDN_2_WPR_note.R"))



``` 


```{r css_js}
# Add standard stylesheets and javascript to support kendo
cat(writeLines(readLines(here("report/resources/headers.htm"))))
```

```{css, echo=FALSE}

/* Table 1.2.1 */
/* Recreating simple striped bootstrap table */
#burden_num_table {
  border-spacing: 0;
  border-collapse: collapse;
  margin-top: 1em;
  margin-bottom: 1em;
  /* Next two lines to allow horizontal scrolling on narrow screens */
  display: block;
  overflow-x: auto;
}

#burden_num_table th {
  border-bottom: 2px solid #DDDDDD;
  padding: 8px;
}

#burden_num_table td {
  border-top: 1px solid #DDDDDD;
  padding: 8px;
}

/* light gray for odd rows */
#burden_num_table tr:nth-child(odd) td {
  background-color: #F5F5F5;	
}

/* Bold for the final row with thick line above */
#burden_num_table tr:last-child td {
  border-top: 2px solid #DDDDDD;
  font-weight:bold;	
}

/* light gray when hovering over a row */
#burden_num_table tr:hover td {
  background-color: #DDDDDD;
}

/* Centre-align all column headings except for the first */
#burden_num_table th:not(:first-child) {
  text-align: center !important;
}

/* prevent numbers from wrapping in any of the columns */
#burden_num_table td {
  white-space: nowrap;
}


```

```{r}
# to format 100 000 as a R object, avoid breaking lines
cm=formatC(1e5, format="f", big.mark=" ", digits=0)

# Estimates year
yr=2024
```



# 1.2 TB mortality

_Draft! Prepared `r Sys.Date()` using country-reported data snapshot files from `r snapshot_date` and incidence estimates from `r est_date`_


`r lnk("Box 1.2.1")` summarizes the methods used to produce estimates of TB mortality between 2010 and `r yr`.

For consistency with international standards (`r ref_lnk("1")`), a clear distinction is made between deaths from TB in people without HIV (classified as deaths caused by TB) and deaths from TB in people with HIV (classified as deaths caused by HIV, with TB as a contributory cause).  

Globally in `r yr`, there were an estimated `r round(global$mort.nh.num[global$year==yr]/1e6,2)` million deaths among HIV-negative people (95% uncertainty interval [UI]: `r round(global$mort.nh.lo.num[global$year==yr]/1e6,2)`–`r round(global$mort.nh.hi.num[global$year==yr]/1e6,2)` million) and an estimated `r thous(round(global$mort.h.num[global$year==yr]/1e3,0))` deaths among people with HIV (95% UI: `r thous(round(global$mort.h.lo.num[global$year==yr]/1e3,0))`–`r thous(round(global$mort.h.hi.num[global$year==yr]/1e3,0))`) (`r lnk("Fig. 1.2.1")`). The combined total was `r round(global$mort.num[global$year==yr]/1e6,2)` million. 


### `r anch("Fig. 1.2.1")`<span class="red">Fig. 1.2.1</span> Global trends in the estimated number of deaths caused by TB (a) and the TB mortality rate (b), 2010&#8211;`r yr`
<div class="subhead">Shaded areas represent 95% uncertainty intervals. The horizontal dashed line shows the `r inc.mort.milestone.yr` milestone of the End TB Strategy, which is a `r (1 - mort.milestone.vs.2015)*100`% reduction in the total number of deaths caused by TB between 2015 and `r inc.mort.milestone.yr`.</div>

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.2.1, fig.alt="Global trends in the estimated number of TB deaths (left) and the mortality rate (right), 2010-2023"}

f1.2.1a_plot <-
  qplot(
    year,
    mort.num / 1e6,
    data = subset(global,year>=2010),
    geom = 'line',
    colour = I('grey20')
  ) +
  geom_ribbon(
    aes(year, ymin = mort.lo.num / 1e6, ymax = mort.hi.num / 1e6),
    fill = I('grey40'),
    alpha = 0.4
  ) +
  ylab('Millions per year') + xlab('') +
  geom_line(aes(year, mort.h.num / 1e6), colour = I('lightskyblue2')) +
  geom_ribbon(
    aes(year, ymin = mort.h.lo.num / 1e6, ymax = mort.h.hi.num / 1e6),
    fill = I('lightskyblue2'),
    alpha = 0.4
  ) +
  geom_line(aes(year, mort.nh.num / 1e6), colour = I('blue')) +
  geom_ribbon(
    aes(year, ymin = mort.nh.lo.num / 1e6, ymax = mort.nh.hi.num / 1e6),
    fill = I('blue'),
    alpha = 0.4
  ) +
  geom_line(aes(year, mort.num / 1e6), colour = I('grey20')) +
  geom_ribbon(
    aes(year, ymin = mort.lo.num / 1e6, ymax = mort.hi.num / 1e6),
    fill = I('grey40'),
    alpha = 0.4
  ) +
  annotate(
    'text',
    y = 0.6,
    x = 2013,
    label = paste(inc.mort.milestone.yr, 'milestone'),
    hjust = 0,
    size = I(4)
  ) +
#  expand_limits(y = 0.1) +
  geom_hline(aes(yintercept = mort.milestone / 1e6), linetype = I(2)) +
  annotate(
    'text',
    x = 2013,
    y = 2,
    label = 'Total',
    hjust = 0,
    size = I(4)
  ) +
  annotate(
    'text',
    x = 2013,
    y = 1.2,
    label = 'HIV-negative people',
    hjust = 0,
    size = I(4)
  ) +
  annotate(
    'text',
    x = 2013,
    y = 0.20,
    label = 'people with HIV',
    hjust = 0,
    size = I(4)
  ) +
 # scale_y_log10() +
  scale_x_continuous(breaks=c(2010,2015,2020)) +
  theme_gtb() + theme(legend.position = 'none')

output_ggplot(f1.2.1a_plot, global[, .(year, mort.num, mort.lo.num, mort.hi.num, mort.nh.num, mort.nh.lo.num, mort.nh.hi.num, mort.h.num, mort.h.lo.num, mort.h.hi.num, mort.milestone)], show_static, pdf_csv_folder, save_csv, save_pdf)

f1.2.1b_plot <-
  qplot(
    year,
    mort,
    data = subset(global,year>=2010),
    geom = 'line',
    colour = I('grey20')
  ) +
  geom_ribbon(aes(year, ymin = mort.lo, ymax = mort.hi),
              fill = I('grey40'),
              alpha = 0.4) +
  ylab('Rate per 100 000 population per year') + xlab('') +
  geom_line(aes(year, mort.h), colour = I('lightskyblue2')) +
  geom_ribbon(
    aes(year, ymin = mort.h.lo, ymax = mort.h.hi),
    fill = I('lightskyblue2'),
    alpha = 0.4
  ) +
  geom_line(aes(year, mort.nh), colour = I('blue')) +
  geom_ribbon(
    aes(year, ymin = mort.nh.lo, ymax = mort.nh.hi),
    fill = I('blue'),
    alpha = 0.4
  ) +
  geom_line(aes(year, mort), colour = I('grey20')) +
  geom_ribbon(aes(year, ymin = mort.lo, ymax = mort.hi),
              fill = I('grey40'),
              alpha = 0.4) +
#  expand_limits(y = 0.1) +
  annotate(
    'text',
    x = 2013,
    y = 28,
    label = 'Total',
    hjust = 0,
    size = I(4)
  ) +
  annotate(
    'text',
    x = 2013,
    y = 15,
    label = 'HIV-negative people',
    hjust = 0,
    size = I(4)
  ) +
  annotate(
    'text',
    x = 2013,
    y = 5,
    label = 'people with HIV',
    hjust = 0,
    size = I(4)
  ) +
 # scale_y_log10() +
  scale_x_continuous(breaks=c(2010,2015,2020)) +
  theme_gtb() + theme(legend.position = 'none')

output_ggplot(f1.2.1b_plot, global[, .(year, mort, mort.lo, mort.hi, mort.h, mort.h.lo, mort.h.hi, mort.nh, mort.nh.lo, mort.nh.hi)], show_static, pdf_csv_folder, save_csv, save_pdf)

```

<div class="col-md-6">
#### (a) Number
<div id="fig_1_2_1a"></div>
</div>
<div class="col-md-6">
#### (b) Rate per `r cm` population
<div id="fig_1_2_1b"></div>
</div>

<hr />
<br />

Following increases in the estimated number of TB deaths during COVID-related disruptions to TB diagnosis and treatment in 2020 and 2021, numbers have subsequently fallen. There was a `r 100-round(100*global$mort.num[global$year==yr]/global$mort.num[global$year==(yr-1)],1) `% reduction between 2023 and 2024 and a `r 100-round(100*global$mort.num[global$year==yr]/global$mort.num[global$year==2021],1) `% reduction compared with 2021. The net reduction in the total number of deaths caused by TB between 2015 and 2024 was `r 100-round(100*global$mort.num[global$year==yr]/global$mort.num[global$year==2015],0)`%, still far from the 2025 milestone of the WHO End TB Strategy (a `r (1 - mort.milestone.vs.2015)*100`% reduction between 2015 and 2025).

In `r yr`, most of the estimated deaths caused by TB among HIV-negative people occurred in the WHO regions of South-East Asia (`r round(100*regional$mort.nh.num[regional$year==yr & regional$g.whoregion=="SEA"]/global$mort.nh.num[global$year==yr],0)`%), Africa (`r round(100*regional$mort.nh.num[regional$year==yr & regional$g.whoregion=="AFR"]/global$mort.nh.num[global$year==yr],0)`%) and the Western Pacific (`r round(100*regional$mort.nh.num[regional$year==yr & regional$g.whoregion=="WPR"]/global$mort.nh.num[global$year==yr],0)`%), with smaller shares in the Eastern Mediterranean (`r round(100*regional$mort.nh.num[regional$year==yr & regional$g.whoregion=="EMR"]/global$mort.nh.num[global$year==yr],1)`%), the Americas (`r round(100*regional$mort.nh.num[regional$year==yr & regional$g.whoregion=="AMR"]/global$mort.nh.num[global$year==yr],1)`%) and Europe (`r round(100*regional$mort.nh.num[regional$year==yr & regional$g.whoregion=="EUR"]/global$mort.nh.num[global$year==yr],1)`%). Most of the TB deaths among people with HIV occurred in the African Region (`r lnk("Table 1.2.1")`).

### `r anch("Table 1.2.1")`<span class="red">Table 1.2.1</span> Global and regional estimates of TB mortality, numbers (in thousands) and rates (per `r cm` population) in `r yr`

<div class="subhead">Low and high are the 5th and 95th percentiles of the uncertainty interval (UI). `r IDN_2_WPR_note2`</div>

```{r tab_1.2.1}

tab.header2 <- c('Region or country group','Best estimate','Low', 'High',
                  'Best estimate', 'Low', 'High',
                  'Best estimate', 'Low', 'High',
                  'Best estimate', 'Low', 'High'
                  )

knitr::kable(cbind(tab1[c(32:37, 31, 38), c(1, 9:14)], tab1b[c(32:37, 31, 38), 8:13]),
             format = "html",
             align = 'rcccccccccccc',
             col.names = tab.header2,
             # Add a table ID so that it can be styled using extra CSS in Sitefinity
             table.attr = "id='burden_num_table'") |>
add_header_above(header = c(" " = 1, "People without HIV" = 3, "People with HIV" = 3, "People without HIV" = 3, "People with HIV" = 3)) |>
add_header_above(header = c(" " = 1, "Number of deaths (in thousands)" = 6, "Rate per 100 000 population" = 6)) 
  # pack_rows("High burden countries", 1, 30) |>
  # pack_rows('Groupings', 31, 38) 

```

<hr />
<br />

```{r}
mort.nh.num= global$mort.nh.num[global$year==yr]
mort.h.num= global$mort.h.num[global$year==yr]
mort.num= global$mort.num[global$year==yr]

mort.nh.num.m=sum(Mglobsplt$mort[Mglobsplt$sex=="M" & Mglobsplt$child==F],na.rm=T)
mort.nh.num.f=sum(Mglobsplt$mort[Mglobsplt$sex=="F" & Mglobsplt$child==F],na.rm=T)
mort.nh.num.c=sum(Mglobsplt$mort[Mglobsplt$child==T],na.rm=T)

mort.h.num.m=sum(MPglobsplt$mort[MPglobsplt$sex=="M" & MPglobsplt$child==F],na.rm=T)
mort.h.num.f=sum(MPglobsplt$mort[MPglobsplt$sex=="F" & MPglobsplt$child==F],na.rm=T)
mort.h.num.c=sum(MPglobsplt$mort[MPglobsplt$child==T],na.rm=T)


```


Globally in `r yr`, `r round(100*mort.nh.num.m/mort.nh.num,1)`% of the HIV-negative people who died from TB were adult men, `r round(100*mort.nh.num.f/mort.nh.num,1)`% were adult women and `r round(100*mort.nh.num.c/mort.nh.num,1)`% were children and young adolescents (aged <15 years) (`r lnk("Fig. 1.2.2")`).

Of the TB deaths among people with HIV, `r round(100*mort.h.num.m/mort.h.num,1)`% were adult men, `r round(100*mort.h.num.f/mort.h.num,1)`% were adult women and `r round(100*mort.h.num.c/mort.h.num,1)`% were children and young adolescents.


### `r anch("Fig. 1.2.2")` <span class="red">Fig. 1.2.2</span> Global distribution of estimated TB mortality in HIV-negative people and in people with HIV by age group and sex (female in <span style="color:#6363C0">purple</span>; male in <span style="color:#F4A81D">orange</span>), `r yr` 

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=6, fig_1.2.2, fig.alt="Main methods used to estimate TB mortality in People without HIV"} 

```

<div class="row">
#### (a) Among HIV-negative people
<div class="col-md-6">
#### Female
<div id="fig_1_2_3a"></div>
</div>
<div class="col-md-6">
#### Male
<div id="fig_1_2_3b"></div>
</div>


#### (b) Among people with HIV
<div class="col-md-6">
#### Female
<div id="fig_1_2_3ha"></div>
</div>
<div class="col-md-6">
#### Male
<div id="fig_1_2_3hb"></div>
</div>

</div>

<hr />
<br />



In `r yr`, an estimated total of `r thous(round(m.tot.male,-3)/1e3)` adult men (aged &#8805;15 years) died from TB (95% UI: `r thous(round(m.tot.male.lo,-3)/1e3)`–`r thous(round(m.tot.male.hi,-3)/1e3)`): `r thous(round(m.nh.male,-3)/1e3)` among those HIV-negative (95% UI: `r thous(round(m.nh.male.lo,-3)/1e3)`–`r thous(round(m.nh.male.hi,-3)/1e3)`) and `r thous(round(m.h.male,-3)/1e3)` among those with HIV (95% UI: `r thous(round(m.h.male.lo,-3)/1e3)`–`r thous(round(m.h.male.hi,-3)/1e3)`). An estimated total of `r thous(round(m.tot.fem,-3)/1e3)` adult women (aged &#8805;15 years) died from TB (95% UI: `r thous(round(m.tot.fem.lo,-3)/1e3)`–`r thous(round(m.tot.fem.hi,-3)/1e3)`): `r thous(round(m.nh.fem,-3)/1e3)` among those HIV-negative (95% UI: `r thous(round(m.nh.fem.lo,-3)/1e3)`–`r thous(round(m.nh.fem.hi,-3)/1e3)`) and `r thous(round(m.h.fem,-3)/1e3)` among those with HIV (95% UI: `r thous(round(m.h.fem.lo,-3)/1e3)`–`r thous(round(m.h.fem.hi,-3)/1e3)`).

In `r yr`, an estimated total of `r thous(round(m.tot.children,-3)/1e3)` children and young adolescents (<15 years) died from TB (95% UI: `r thous(round(m.tot.children.lo,-3)/1e3)`–`r thous(round(m.tot.children.hi,-3)/1e3)`): `r thous(round(m.nh.tot,-3)/1e3)` among those HIV-negative (95% UI: `r thous(round(m.nh.tot.lo,-3)/1e3)`–`r thous(round(m.nh.tot.hi,-3)/1e3)`) and `r hund(m.h.tot)` among those with HIV (95% UI: `r hund(m.h.tot.lo)`–`r hund(m.h.tot.hi)`). Among HIV-negative children and young adolescents, the estimated number of TB deaths in `r yr` was `r thous(round(m.nh.l5,-3)/1e3)` among children aged <5 years (95% UI: `r thous(round(m.nh.l5.lo,-3)/1e3)`–`r thous(round(m.nh.l5.hi,-3)/1e3)`); and `r thous(round(m.nh.5.14,-3)/1e3)` in older children and young adolescents aged 5–14 years (95% UI: `r thous(round(m.nh.5.14.lo,-3)/1e3)`–`r thous(round(m.nh.5.14.hi,-3)/1e3)`).

The latest year for which WHO has published estimates of global deaths by cause is 2021. In that year, TB was the tenth leading cause of death worldwide (`r lnk("Fig. 1.2.3")`) and the second leading cause of death from a single infectious agent, after COVID-19. In the African and South-East Asia regions, it was the fourth and fifth leading cause of death, respectively. 

In 2024, a total of `r thous(70)` deaths from COVID-19 were officially reported to WHO (`r ref_lnk("2")`). This number does not account for late reporting or underreporting, but it is far below the estimated number of deaths from TB.

### `r anch("Fig. 1.2.3")`<span class="red">Fig. 1.2.3</span> Top 15 causes of death worldwide in 2021 ^a,b^
<div class="subhead">Deaths from TB among people with HIV are shown in grey.</div>

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=6, fig_1.2.3, fig.alt="Top causes of death worldwide in 2021"} 

tbhiv <- global$mort.h.num[global$year == 2021] / 1e3
top10$tbhiv <- c(rep(0, 12), tbhiv, rep(0, 7))
top10$n <- top10$deaths / 1000 + top10$tbhiv / 1000

top2021$tbhiv <- c(rep(0, 9), tbhiv, rep(0, 5))
top2021$n <- top2021$deaths / 1000 + top2021$tbhiv
top2021$tbhiv2 <- 1000*top2021$tbhiv


f1.2.4_plot <- ggplot(data = top2021, aes(reorder(causename, deaths), n)) +
  geom_bar(stat = 'identity',
           fill = I('grey70'),
           colour = I('black')) +
  geom_bar(
    aes(causename, deaths / 1000),
    stat = 'identity',
    fill = I('#00aaad'),
    colour = I('black')
  ) +
  xlab('')  + ylab('Number of deaths (millions)') +
  scale_y_continuous(breaks = 0:9) +
  coord_flip() +
  theme_gtb()

output_ggplot(f1.2.4_plot, top2021, show_static, pdf_csv_folder, save_csv, save_pdf)

```
<div id="fig_1_2_4"></div>
<div class="footnote"> ^a^ This is the latest year for which estimates for all causes are currently available. See WHO estimates, available at https://www.who.int/data/gho/data/themes/mortality-and-global-health-estimates/ghe-leading-causes-of-death.<br>
^b^ Deaths from TB among people with HIV are officially classified as deaths caused by HIV/AIDS in the International Classification of Diseases. </div> 

<hr />
<br />

The estimated number of deaths officially classified as caused by TB (i.e. those among HIV-negative people) in `r yr`, was almost double the 630 000 (95% UI: 600 000–660 000) caused by HIV/AIDS (`r lnk("Fig. 1.2.4")`).  

### `r anch("Fig. 1.2.4")`<span class="red">Fig. 1.2.4</span> Estimated number of deaths caused by HIV/AIDS and TB in `r yr` ^a,b^ 
<div class="subhead">Deaths from TB among people with HIV are shown in grey.</div>

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png',fig.height=6, fig_1.2.4, fig.alt="Estimated number of deaths from HIV/AIDS and TB in 2023"} 

hivd <- unaids[, .(sum(mort.hiv.num, na.rm = T)), by = year]
cod <- data.table(
  cause = c('HIV/AIDS', 'TB'),
  n = c((last(hivd$V1) - last(global$mort.h.num)) / 1e6, last(global$mort.nh.num) /
          1e6),
  tbhiv = rep(last(global$mort.h.num) / 1e6, 2)
)
cod[, total := n + tbhiv]

f1.2.5_plot <- ggplot(data = cod, aes(reorder(cause, total), total)) +
  geom_bar(stat = 'identity',
           fill = I('grey70'),
           colour = I('black')) +
  geom_bar(
    aes(cause, n),
    stat = 'identity',
    fill = I('#00aaad'),
    colour = I('black')
  ) +
  xlab('') + ylab('Millions (2023)') +
  coord_flip() +
  theme_gtb()

output_ggplot(f1.2.5_plot, cod, show_static, pdf_csv_folder, save_csv, save_pdf)

```
<div id="fig_1_2_5"></div>
<div class="footnote"> ^a^ For HIV/AIDS, the latest estimates of the number of deaths in `r yr` are those published by UNAIDS (`r ref_lnk("3")`). For TB, the estimates for `r yr` are those published in this report.<br>
^b^ Deaths from TB among people with HIV are officially classified as deaths caused by HIV/AIDS in the International Classification of Diseases.</div> 

<hr />
<br />

Since 2010, deaths from TB among HIV-negative people have fallen much more slowly than deaths caused by HIV and were also more impacted by the COVID-19 pandemic (`r lnk("Fig. 1.2.5")`).

### `r anch("Fig. 1.2.5")`<span class="red">Fig. 1.2.5</span> Global trends in the estimated number of deaths caused by TB and HIV (in millions), 2010&#8211;`r yr` ^a,b^
<div class="subhead">Shaded areas represent 95% uncertainty intervals.</div> 

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.2.5, fig.alt="Global trends in the estimated number of deaths caused by TB and HIV (in millions), 2000–2023"} 

m <- 1e5
ghiv.mort <-
  est[, addXY(mort.hiv / m, r.sd = mort.hiv.sd / m, weights = as.numeric(pop)), by =
        year]
dta <- merge(global, ghiv.mort, by = 'year')
M <- 1e6

f1.2.6_plot <- ggplot(data = subset(dta,year>=2010), aes(year, mort.h.num / M)) +
  geom_line(colour = I('lightskyblue2')) +
  geom_line(aes(year, mort.nh.num / M), colour = I('blue')) +
  geom_ribbon(
    aes(year, ymin = mort.h.lo.num / M, ymax = mort.h.hi.num / M),
    fill = I("lightskyblue2"),
    alpha = I(.4)
  ) +
  geom_ribbon(
    aes(year, ymin = mort.nh.lo.num / M, ymax = mort.nh.hi.num / M),
    fill = I("blue"),
    alpha = I(.4)
  ) +
  geom_line(aes(year, r.num / M), colour = I('grey50')) +
  geom_ribbon(
    aes(year, ymin = r.lo.num / M, ymax = r.hi.num / M),
    fill = I("grey50"),
    alpha = I(.4)
  ) +
  xlab('') + ylab('Millions of deaths per year') +
  annotate(
    'text',
    x = 2013,
    y = 0.8,
    label = 'HIV deaths',
    size = I(4),
    colour = I('grey50')
  ) +
  annotate(
    'text',
    x = 2013,
    y = 1.6,
    label = 'TB deaths in\n HIV-negative people',
    size = I(4),
    colour = I('blue')
  ) +
  annotate(
    'text',
    x = 2013,
    y = 0.2,
    label = 'TB deaths in\npeople with HIV',
    size = I(4),
    colour = I('lightskyblue4')
  ) +
  expand_limits(y = c(0.2,1.8)) + 
  scale_y_continuous(breaks=seq(0.2,1.8,0.2)) +
  scale_x_continuous(breaks=c(2010,2015,2020)) +
    theme_gtb()


output_ggplot(f1.2.6_plot, dta[, .(year, mort.h.num, mort.h.lo.num, mort.h.hi.num, mort.nh.num, mort.nh.lo.num, mort.nh.hi.num, r.num, r.lo.num, r.hi.num)], show_static, pdf_csv_folder, save_csv, save_pdf)

```
<div id="fig_1_2_6"></div>
<div class="footnote"> ^a^ For HIV/AIDS, the latest estimates of the number of deaths in `r yr` are those published by UNAIDS (`r ref_lnk("3")`). For TB, the estimates for `r yr` are those published in this report.<br>
^b^ Deaths from TB among people with HIV are officially classified as deaths caused by HIV/AIDS in the International Classification of Diseases.</div> 

<hr />
<br />


The TB mortality rate fell in five of the six WHO regions in `r yr` and stabilized in the Region of the Americas (`r lnk("Fig. 1.2.6")`).

### `r anch("Fig. 1.2.6")`<span class="red">Fig. 1.2.6</span> Trends in estimated TB mortality rates by WHO region, 2010&#8211;`r yr`
<div class="subhead">Estimated TB mortality rates among HIV-negative people are shown in <span style="color:#3232ff">blue</span> and estimated mortality rates among people with HIV are shown in <span style="color:#87ceeb">light blue</span>. Shaded areas represent 95% uncertainty intervals. `r IDN_2_WPR_note2`</div> 


```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height = 8, fig_1.2.6, fig.alt="Trends in estimated TB mortality rates by WHO region, 2010−2023"} 

f1.2.7_plot <- qplot(
    year,
    mort.nh,
    data = subset(regional,year>=2010),
    geom = 'line',
    colour = I('blue')
  ) +
  geom_ribbon(
    aes(year, ymin = mort.nh.lo, ymax = mort.nh.hi),
    fill = I('blue'),
    alpha = 0.4
  ) +
  geom_line(aes(year, mort.h), colour = I('lightskyblue2')) +
  geom_ribbon(
    aes(year, ymin = mort.h.lo, ymax = mort.h.hi),
    fill = I('lightskyblue2'),
    alpha = 0.4
  ) +
  expand_limits(y=0.1) +
  facet_wrap(~ region, scale = 'free_y') +
  xlab('') + ylab('Mortality rate per 100 000 population per year') +
  #  scale_y_continuous(trans='log10', breaks=c(0.1,1,5,10,20,30,50)) +
  theme_gtb() +
  #scale_y_log10() +
  scale_x_continuous(breaks=c(2010,2015,2020)) +
  theme(legend.position = 'none')

output_ggplot(f1.2.7_plot, regional[, .(region, year,  mort.nh,  mort.nh.lo,  mort.nh.hi, mort.h, mort.h.lo, mort.h.hi)], show_static, pdf_csv_folder, save_csv, save_pdf)

```
<div class="row">
<div class="col-md-4">
<div id="fig_1_2_7_AFR"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_7_AMR"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_7_SEAR"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_2_7_EUR"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_7_EMR"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_7_WPR"></div>
</div>
</div>

<hr />
<br />


Trends in the absolute number of TB deaths at regional level vary (`r lnk("Fig. 1.2.7")`). The pattern of a fall in the absolute number of deaths caused by TB (including those among people with HIV) until 2019, followed by increases during the COVID-19 pandemic and then declines starting in either 2022 or 2023, is evident in the WHO European, South-East Asia and Western Pacific regions. 

In the Eastern Mediterranean Region, an increase in 2020 and 2021 was followed by a decline in 2022 and then by a small increase between 2022 and `r yr`. In the Region of the Americas, the estimated number of deaths caused by TB peaked in 2022, declined in 2023 and stabilized in `r yr`. In the African Region, the estimated number of deaths caused by TB has fallen year on year since 2011.

### `r anch("Fig. 1.2.7")`<span class="red">Fig. 1.2.7</span> Trends in the estimated absolute number of TB deaths in HIV-negative people and people with HIV (in thousands) by WHO region, 2010&#8211;`r yr` 
<div class="subhead">Shaded areas represent 95% uncertainty intervals. The horizontal dashed line shows the `r inc.mort.milestone.yr` milestone of the End TB Strategy, which is a `r (1 - mort.milestone.vs.2015)*100`% reduction in the total number of deaths caused by TB between 2015 and `r inc.mort.milestone.yr`. `r IDN_2_WPR_note2`</div> 


```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height = 8, fig_1.2.7, fig.alt="Trends in the estimated absolute number of TB deaths (people with HIV and People without HIV, in thousands) by WHO region, 2010-2023"} 

f1.2.8_plot <- qplot(
    year,
    mort.num / 1e3,
    data = subset(regional,year>=2010),
    geom = 'line',
    colour = I('black')
  ) +
  geom_ribbon(
    aes(year, ymin = mort.lo.num / 1e3, ymax = mort.hi.num / 1e3),
    fill = I('black'),
    alpha = 0.2
  ) +
  geom_hline(aes(yintercept = mort.milestone / 1e3), linetype = I(2)) +
#  expand_limits(y=0.1) +
  facet_wrap(~ region, scale = 'free_y') +
  xlab('') + ylab('Thousands per year') +
  #scale_y_log10() +
  theme_gtb() +
  theme(legend.position = 'none')


output_ggplot(f1.2.8_plot, regional[, .(region,year,mort.num,mort.lo.num,mort.hi.num,mort.milestone)], show_static, pdf_csv_folder, save_csv, save_pdf)

```
<div class="row">
<div class="col-md-4">
<div id="fig_1_2_8_AFR"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_8_AMR"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_8_SEAR"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_2_8_EUR"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_8_EMR"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_8_WPR"></div>
</div>
</div>

<hr />
<br />

The European Region and the African Region are the closest to reaching the `r inc.mort.milestone.yr` milestone of the End TB Strategy, with reductions of `r round(100-100*regional$mort.num[regional$g.whoregion=="EUR" & regional$year==yr]/regional$mort.num[regional$g.whoregion=="EUR" & regional$year==2015],0)`% and `r round(100-100*regional$mort.num[regional$g.whoregion=="AFR" & regional$year==yr]/regional$mort.num[regional$g.whoregion=="AFR" & regional$year==2015],0)`% respectively between 2015 and `r yr`. The four other regions are still far from the `r inc.mort.milestone.yr` milestone of the End TB Strategy, with progress up to 2019 set back by disruptions during the COVID-19 pandemic and its aftermath.

Trends in TB deaths by income group are shown in (`r lnk("Fig. 1.2.8")`).

### `r anch("Fig. 1.2.8")`<span class="red">Fig. 1.2.8</span> Trends in the estimated absolute number of TB deaths in HIV-negative people and people with HIV (in thousands) by income group, 2010&#8211;`r yr`
<div class="subhead">Shaded areas represent 95% uncertainty intervals. The horizontal dashed line shows the `r inc.mort.milestone.yr` milestone of the End TB Strategy, which is a `r (1 - mort.milestone.vs.2015)*100`% reduction in the total number of deaths caused by TB between 2015 and `r inc.mort.milestone.yr`.</div> 

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=8,  fig_1.2.8, fig.alt="Trends in the estimated absolute number of TB deaths (HIV-negative people and people with HIV, in thousands) by income group, 2000-2023"} 




f1.2.9 <- qplot(
    year,
    mort.num / 1e3,
    data = subset(income,year>=2010),
    geom = 'line',
    colour = I('black')
  ) +
  geom_ribbon(
    aes(year, ymin = mort.lo.num / 1e3, ymax = mort.hi.num / 1e3),
    fill = I('black'),
    alpha = 0.2
  ) +
  geom_hline(aes(yintercept = mort.milestone/ 1e3), linetype = I(2)) +

  #  expand_limits(y=0.1) +
  facet_wrap(~ group, scale = 'free_y') +
  xlab('') + ylab('Total TB deaths (thousands)') +
  #scale_y_log10() +
  theme_gtb() +
  theme(legend.position = 'none')


f1.2.9_data=income[, .(group,year,mort.num,mort.lo.num,mort.hi.num,mort.milestone)]

output_ggplot(f1.2.9, f1.2.9_data, show_static, pdf_csv_folder, save_csv, save_pdf)


```


<div class="row">
<div class="col-md-6">
<div id="fig_1_2_8a_LIC"></div>
<div id="fig_1_2_8a_LMC"></div>
</div>

<div class="col-md-6">
<div id="fig_1_2_8a_UMC"></div>
<div id="fig_1_2_8a_HIC"></div>
</div>
</div>

<hr />
<br />




The regional distribution of TB deaths by age and sex in `r yr` varies by WHO region (`r lnk("Fig. 1.2.9")`). 

### `r anch("Fig. 1.2.9")`<span class="red">Fig. 1.2.9</span> Regional distribution of estimated TB mortality in HIV-negative people and people with HIV by age group and sex (female in <span style="color:#6363C0">purple</span>; male in <span style="color:#F4A81D">orange</span>), `r yr`
<div class="subhead">`r IDN_2_WPR_note2`</div>



```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=10, fig_1.2.9a, fig.alt="Main methods used to estimate TB mortality in people without HIV"} 
# 
# disag.plot.mort.regional <- function(D){
#     ## ages and colors etc
#     agz3 <- c('0\u20134','5\u201314','15\u201324','25\u201334','35\u201344','45\u201354',
#               '55-64','65plus') #for labels
#     agz4 <- c(rev(rev(agz3)[-1]),"\u226565")
#     # whozt <- c('Africa','The Americas','Eastern Mediterranean','Europe',
#     #            'South-East Asia',
#     #            'Western Pacific')
#     # regs <- c('AFR','AMR','EMR','EUR','SEA','WPR')
#     clz <- c('F'=palette_gtb('female'),
#              'M'=palette_gtb('male'))
#     wr <- levels(D$name)
#     ## plot construction
#     Pltlst <- list()
#     for(i in 1:6){
# #        plt <- muplot(D[g.whoregion==regs[i]],whozt[i]) +
#         plt <- muplot(D[name==wr[i]],wr[i]) +
#             theme(legend.position="none")
# #        if(!(i%%3==1)) plt <- plt + theme(axis.text.y=element_blank())
#         plt <- plt + coord_flip() +
#             scale_fill_manual(values=clz)+
#             theme_gtb()+
#             theme(axis.title.x=element_blank(),
#                   axis.text.x=element_blank(),
#                   axis.ticks.x=element_blank(),
#                   legend.position = 'none')
#         Pltlst[[i]] <- plt
#     }
#     p <- ggpubr::ggarrange(plotlist = Pltlst, ncol = 3,nrow=2)
#     annotate_figure(p, left = text_grob('Age group (years)', rot=90))
# }
# 
# disag.plot.mort.regional(Mregsplt)

```

<div class="row">

<div class="col-md-6">
#### (a) Female
<div id="fig_1_2_9a_AFR"></div>
<div id="fig_1_2_9a_AMR"></div>
<div id="fig_1_2_9a_EMR"></div>
<div id="fig_1_2_9a_EUR"></div>
<div id="fig_1_2_9a_SEA"></div>
<div id="fig_1_2_9a_WPR"></div>
</div>
	
<div class="col-md-6">
#### (b) Male
<div id="fig_1_2_9b_AFR"></div>
<div id="fig_1_2_9b_AMR"></div>
<div id="fig_1_2_9b_EMR"></div>
<div id="fig_1_2_9b_EUR"></div>
<div id="fig_1_2_9b_SEA"></div>
<div id="fig_1_2_9b_WPR"></div>
</div>
	
</div>


<hr />
<br />


```{r}
dta <- subset(est, year==yr, select=c("iso3","country","year","mort.num"))
dta2 <- subset(est, year==2015, select=c("iso3","year","mort.num"))
dta2$mort.num.2021=dta2$mort.num
dta2$mort.num=NULL

dta=merge(dta,dta2,by="iso3")
dta$rel.red.mort.num=(1-(dta$mort.num/dta$mort.num.2021))*100
dta$rel.red.mort.num.cat=ifelse(dta$rel.red.mort.num>0,"Reduction","Increase")

dta$rel.red.mort.num.fordisplay=ifelse(dta$rel.red.mort.num<0,
                                      -round(dta$rel.red.mort.num,1),
                                      round(dta$rel.red.mort.num,1))

dta$var <- cut(
  dta$rel.red.mort.num,
  c(-Inf,-5,5, 20, 35, 50,Inf),
  c('Increase (>5%)',"Stable (-5% to +5%)", 'Decrease (6\u201319%)', 'Decrease (20\u201334%)', 'Decrease (35\u201349%)','Decrease (\u226550%)'),
  right = F,
  ordered_result = T
)

dta=merge(dta,cty[,.(iso3,g.whostatus)],by="iso3")
reach.milestone.lst=dta[(var=="Decrease (\u226550%)" | var=="Decrease (35\u201349%)") & g.whostatus=="M",unique(iso3)]
n.reach.milestone=length(reach.milestone.lst)


```



Progress in reducing the total number of deaths caused by TB (among people with and without HIV) at country level is highly variable (`r lnk("Fig. 1.2.10")`). A total of `r n.reach.milestone` countries have surpassed the first (2020) milestone of the End TB Strategy, which was a 35% reduction compared with the baseline of 2015. At the other extreme, the number of deaths caused by TB has increased in some countries, mainly related to disruptions during the COVID-19 pandemic. This increase has been particularly pronounced in the Region of the Americas.

### `r anch("Fig. 1.2.10")`<span class="red">Fig. 1.2.10</span> Change (%) in the estimated number of deaths caused by TB among HIV-negative people and people with HIV, `r yr` compared with 2015

<div class="subhead"> The first milestone of the End TB Strategy was a 35% reduction by 2020, compared with 2015; the second milestone is a 75% reduction by 2025, compared with 2015. The last two categories (decrease 35&#8211;49%, and decrease &#8805;50%) distinguish the countries that have made the most progress towards the second milestone of the End TB Strategy.</div>

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.2.10, fig.alt="Change (%) in the estimated number of deaths caused by TB, 2023 compared with 2015"}

if (save_hidef == TRUE) {
  # Generate a high definition map for the PDF
  # Update fig numbering according to the masterlist of figs/tables
  f14_data <- dta
  
  f14_plot <- whomapper(
  X = f14_data, water_col = 'white',
  legend_title = "",
  legend_pos = "none",
  colours = c("lightyellow2","lightyellow3",brewer.pal(5, "Blues")[2:5]),
  hidef = save_hidef
)

output_ggplot(f14_plot, f14_data, show_static, pdf_csv_folder, save_csv, save_pdf)

} else {
  # Generate a standard map for the web page

# f1.2.11_plot <- whomapper(
#   X = dta, water_col = 'white',
#   legend_title = "Change (2024 vs 2015, %)",
#   legend_pos = c(0.14, 0.34),
#   colours = c("lightyellow2","lightyellow3",brewer.pal(4, "Blues"))
# )

f1.2.11_plot <- whomapper(
  X = dta, water_col = 'white',
  legend_title = "Change (2024 vs 2015, %)",
  legend_pos = c(0.14, 0.34),
  colours = c("lightyellow2","lightyellow3",brewer.pal(5, "Blues")[2:5])
)


output_ggplot(f1.2.11_plot, dta, show_static=TRUE, pdf_csv_folder, save_csv, save_pdf)

}

```

<hr />
<br />

Trends  in the total number of deaths caused by TB (among people with and without HIV) in the 30 high TB burden countries are also highly variable. Between 2019 and 2021, striking increases are estimated to have occurred in countries that experienced major shortfalls in TB notifications in 2020 and 2021 (e.g. Indonesia, Myanmar, the Philippines), while in others previous declines have slowed or stabilized.

In `r yr`, the best estimate of the total number of deaths caused by TB suggested that the 2025 milestone of the WHO End TB Strategy (a `r (1 - mort.milestone.vs.2015)*100`% reduction by 2025, compared with 2015) was close to being reached in five high TB burden countries: the Central African Republic, Kenya, Nigeria, the United Republic of Tanzania and Zambia (`r lnk("Fig. 1.2.11")`). 

### `r anch("Fig. 1.2.11")`<span class="red">Fig. 1.2.11</span> Trends in the estimated absolute number (in thousands) of deaths caused by TB among HIV-negative people and people with HIV in the 30 high TB burden countries, 2010&#8211;`r yr`
<div class="subhead">Shaded areas represent 95% uncertainty intervals. The horizontal dashed line shows the `r inc.mort.milestone.yr` milestone of the End TB Strategy, which is a `r (1 - mort.milestone.vs.2015)*100`% reduction in the total number of deaths caused by TB between 2015 and `r inc.mort.milestone.yr`.</div> 

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=12, fig_1.2.11, fig.alt="Trends in the estimated absolute number of TB deaths (people with HIV and People without HIV) in the 30 high TB burden countries, 2010−2023"} 

f1.2.12_plot <-  qplot(
    year,
    mort.num / 1e3,
    data = subset(hest,year>=2010),
    geom = 'line',
    colour = I('grey20')
  ) +
  geom_ribbon(
    aes(year, ymin = mort.lo.num / 1e3, ymax = mort.hi.num / 1e3),
    fill = I('black'),
    alpha = 0.2
  ) +
  geom_hline(aes(yintercept = mort.milestone / 1e3), linetype = I(2)) +
  #facet_wrap(~ country, scales = 'free_y', ncol = 5) +
    facet_wrap(~country, nrow=6, scales="free",
                     # Use the labeller function to make sure long country names are wrapped in panel headers
                     labeller = label_wrap_gen(width = 25)) +
  xlab('') + ylab('TB deaths (total, in thousands) per year') +
#  expand_limits(y = 0.1) +
  #scale_y_log10() +
  scale_x_continuous(breaks=c(2010,2015,2020)) +
  theme_gtb()

output_ggplot(f1.2.12_plot, hest[, .(country,year,mort.num,mort.lo.num,mort.hi.num,mort.milestone)], show_static, pdf_csv_folder, save_csv, save_pdf)

```
<div class="row">
<div class="col-md-4">
<div id="fig_1_2_11_AGO"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_BGD"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_BRA"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_2_11_CAF"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_CHN"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_COG"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_2_11_PRK"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_COD"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_ETH"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_2_11_GAB"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_IND"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_IDN"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_2_11_KEN"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_LSO"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_LBR"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_2_11_MNG"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_MOZ"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_MMR"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_2_11_NAM"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_NGA"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_PAK"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_2_11_PNG"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_PHL"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_SLE"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_2_11_ZAF"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_THA"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_UGA"></div>
</div>
</div>

<div class="row">
<div class="col-md-4">
<div id="fig_1_2_11_TZA"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_VNM"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_11_ZMB"></div>
</div>
</div>
<div class="footnote">Mortality estimates are not shown for the Democratic People’s Republic of Korea, since they are currently under review.</div>

<hr />
<br />

Of the three global TB watchlist countries, the Russian Federation is well on the way to reaching the 2025 milestone (`r lnk("Fig. 1.2.12")`), with a net reduction of `r round(100-100*est$mort.num[est$iso3=="RUS" & est$year==yr]/est$mort.num[est$iso3=="RUS" & est$year==2015],0)`% between 2015 and `r yr`. 

### `r anch("Fig. 1.2.12")`<span class="red">Fig. 1.2.12</span> Trends in the estimated absolute number (in thousands) of deaths caused by TB among HIV-negative people and people with HIV in the three global TB watchlist countries, 2010&#8211;`r yr`
<div class="subhead">Shaded areas represent 95% uncertainty intervals. The horizontal dashed line shows the `r inc.mort.milestone.yr` milestone of the End TB Strategy, which is a `r (1 - mort.milestone.vs.2015)*100`% reduction in the total number of deaths caused by TB between 2015 and `r inc.mort.milestone.yr`.</div> 

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig.height=2.7, fig.width=8, fig.align='left', fig_1.2.12, fig.alt="Trends in the estimated absolute number of TB deaths (people with HIV and HIV-negative people) in the 3 WHO watch-list countries, 2010−2023"} 

dta <- est[iso3 %in% c('KHM','RUS','ZWE')]
dta[, mort.milestone := mort.num[16] * mort.milestone.vs.2015, by = iso3]

f1.2.13_plot <- qplot(
    year,
    mort.num / 1e3,
    data = subset(dta,year>=2010),
    geom = 'line',
    colour = I('grey20')
  ) +
  geom_ribbon(
    aes(year, ymin = mort.lo.num / 1e3, ymax = mort.hi.num / 1e3),
    fill = I('blue'),
    alpha = 0.4
  ) +
  geom_hline(aes(yintercept = mort.milestone / 1e3), linetype = I(2)) +
  #facet_wrap(~ country, scales = 'free_y', ncol = 5) +
    facet_wrap(~country, nrow=1, scales="free",
                     # Use the labeller function to make sure long country names are wrapped in panel headers
                     labeller = label_wrap_gen(width = 25)) +
  xlab('') + ylab('TB deaths (total, in thousands)\nper year') +
  scale_x_continuous(breaks=c(2010,2015,2020)) +
#  expand_limits(y = 0.1) +
  #scale_y_log10() +
  theme_gtb()

output_ggplot(f1.2.13_plot, dta[, .(country,year,mort.num,mort.lo.num,mort.hi.num,mort.milestone)], show_static, pdf_csv_folder, save_csv, save_pdf)


```
<div class="row">
<div class="col-md-4">
<div id="fig_1_2_12_KHM"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_12_RUS"></div>
</div>
<div class="col-md-4">
<div id="fig_1_2_12_ZWE"></div>
</div>
</div>

<hr />
<br />

There is considerable national variation in the TB mortality rate; in `r yr`, most of the countries with the highest rates were in the African Region (`r lnk("Fig. 1.2.13")`). 

### `r anch("Fig. 1.2.13")`<span class="red">Fig. 1.2.13</span> Estimated TB mortality rates in HIV-negative people, `r yr`

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.2.13, fig.alt="Estimated TB mortality rates in HIV-negative people, 2023"} 

dta <- est[year == yr]
dta$var <- cut(
  dta$mort.nh,
  c(0, 1, 5, 20, 40, Inf),
  c('0\u20130.9', '1\u20134.9', '5\u201319', '20\u201339', '\u226540'),
  right = F,
  ordered_result = T
)

f1.2.14_plot <- whomapper(
  X = dta, water_col = 'white',
  legend_title = "Mortality\nper 100 000\npopulation per year",
  legend_pos = c(0.14, 0.34),
  colours = brewer.pal(5, "RdPu"),
)

output_ggplot(f1.2.14_plot, dta[, .(country, mort.nh, var)], show_static=TRUE, pdf_csv_folder, save_csv, save_pdf)

```

<hr />
<br />

There is also considerable variation in the case fatality rate (CFR) i.e. the estimated proportion of people who develop TB that die from the disease (`r lnk("Fig. 1.2.14")`). Globally, the CFR was `r round(100*global$cfr[global$year==yr],1)`% in `r yr`, down from `r round(100*global$cfr[global$year==2023],1)`% in 2023 and `r round(100*global$cfr[global$year==2022],1)`% in 2022 (and below the pre-pandemic level of `r round(100*global$cfr[global$year==2019],1)`% in 2019). 

### `r anch("Fig. 1.2.14")`<span class="red">Fig. 1.2.14</span> Estimates of the case fatality rate (CFR), including HIV-negative people and people with HIV, `r yr`

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.2.14, fig.alt="Case fatality ratio (2023)"} 

dta <-
  est[year == yr & inc > 0, .(iso3,
                        cfr = 100 * mort / inc)]
dta$var <- cut(dta$cfr, 
               breaks=c(0, 5, 10, 15, 20, Inf),
               labels=c('0\u20134.9','5\u20139.9','10\u201314','15\u201319','\u226520'))

f1.2.15_plot <- whomapper(
  X = dta, water_col = 'white',
  colours = brewer.pal(5, "YlOrRd"),
  legend_pos = c(0.14, 0.34),
  legend_title = 'Case fatality rate (%)',
)

output_ggplot(f1.2.15_plot, dta, show_static=TRUE, pdf_csv_folder, save_csv, save_pdf)

```

`r anch("Box 1.2.1")`


```{r}

load(here("inc_mort/analysis/est.method.rda"))

est.meth.mort=est.method[,.(iso3,method.mort)]

est.meth.mort= est.meth.mort %>% mutate(
  var=case_when(
  method.mort>=1 & method.mort<=3 ~ "CRVS data",
  method.mort>=4 & method.mort<=6 ~ "IHME estimates",
  method.mort>=7 & method.mort<=9 ~ "CFR-based",
  method.mort==99 ~ "Country-specific customization"
)
)

cm.lst=unique(est.meth.mort$iso3[est.meth.mort$method.mort==2 | est.meth.mort$method.mort==5 |
                                est.meth.mort$method.mort==8])

rm.lst=unique(est.meth.mort$iso3[est.meth.mort$method.mort==3 | est.meth.mort$method.mort==6 |
                                est.meth.mort$method.mort==9])

n.crvs=sum(est.meth.mort$var=="CRVS data",na.rm=T)
n.ihme=sum(est.meth.mort$var=="IHME estimates",na.rm=T)
n.cfr=sum(est.meth.mort$var=="CFR-based",na.rm=T)



# Share in 2024
#est19=merge(est[year==2019,.(iso3,mort.num)],est.meth.mort[,.(iso3,var)],by="iso3")
est24=merge(est.full[year==2024,.(iso3,mort.num)],est.meth.mort[,.(iso3,var)],by="iso3")

#m19=global$mort.num[global$year==2019]
m24=global$mort.num[global$year==2024]

share.crvs=round(100*sum(est24$mort.num[est24$var=="CRVS data"],na.rm=T)/m24,0)

share.ihme=round(100*sum(est24$mort.num[est24$var=="IHME estimates"],na.rm=T)/m24,0)

share.cfr=round(100*sum(est24$mort.num[est24$var=="CFR-based"],na.rm=T)/m24,0)

share.cm=round(100*sum(est24$mort.num[est24$iso3 %in% cm.lst],na.rm=T)/m24,0)
share.rm=round(100*sum(est24$mort.num[est24$iso3 %in% rm.lst],na.rm=T)/m24,1)

n.mort.mod=length(c(cm.lst,rm.lst))



```



<div class="textbox">
## Box 1.2.1 

## Methods used by WHO to estimate TB mortality 

The main methods used by WHO to estimate TB mortality at country level in the period 2010–`r yr` are shown in  `r lnk("Fig. 1.2.15")`. These methods adhere to global guidelines for accurate and transparent reporting of health estimates (`r ref_lnk("4")`) and are described in detail in a <span class="red">technical appendix</span>.

The methods used for the period 2010–2024 can be summarized as follows: 

* For `r n.crvs` countries, estimates of the number of deaths caused by TB among HIV-negative people were based on data on causes of death from national or sample vital registration (VR) systems that have been officially reported to WHO. For India, officially published data from the sample registration system (SRS) were used (`r ref_lnk("5")`); detailed data from the SRS have not been officially reported to WHO. These `r n.crvs` countries  collectively accounted for `r share.crvs`% of the estimated global number of deaths (among HIV-negative people) caused by TB in 2024.

* For `r n.ihme` countries, estimates published by the Institute of Health Metrics and Evaluation (IHME) were used (`r ref_lnk("6")`). These are based on analyses of VR data as well as additional data. These `r n.ihme`  countries collectively accounted for `r share.ihme`% of the estimated global number of deaths (among HIV-negative people) caused by TB in 2024.

* For the remaining `r n.cfr` countries, TB mortality among HIV-negative people was estimated indirectly as the product of TB incidence and the CFR. These `r n.cfr` countries collectively accounted for `r share.cfr`% of the estimated global number of deaths (among HIV-negative people) caused by TB in 2024.

For all countries, TB mortality among people with HIV was estimated as the product of TB incidence and the CFR, with the latter accounting for the protective effect of antiretroviral treatment (ART).

To produce estimates of TB mortality in the period 2020–`r yr`, country or region-specific dynamic models were developed for the `r n.mort.mod` countries with the biggest absolute reductions in TB notifications during the COVID-19 pandemic (when these reductions departed from pre-2020 trends).

For all other countries and areas, estimates for 2020–`r yr` were based on the same methods as those used up to 2019. 

The <span style="color:#f4a81d">orange</span> dots show countries for which a country-specific dynamic model has been used (n=`r length(cm.lst)`) and the <span style="color:#754d06">brown</span> dots show countries for which a region-specific dynamic model has been used (n=`r length(rm.lst)`) . These two groups of countries accounted for `r share.cm`% and `r share.rm`% of the estimated total number of deaths caused by TB in `r yr`, respectively.


### `r anch("Fig. 1.2.15")`<span class="red">Fig. 1.2.15</span> Main methods used to estimate TB mortality in HIV-negative people, 2010&#8211;`r yr`
<div class="subhead">The <span style="color:#f4a81d">orange</span> dots on the map show the countries for which a country-specific dynamic model has been used and the  <span style="color:#754d06">brown</span> dots show the countries for which a region-specific dynamic model has been used.</div> 

```{r echo=FALSE, message=FALSE, warning=FALSE, results = "asis", dev = 'png', fig_1.2.15, fig.alt="Main methods used to estimate TB mortality in people without HIV, 2000-2024"} 



load(here("inc_mort/analysis/est.method.rda"))

est.meth.mort=est.method[,.(iso3,method.mort)]

est.meth.mort= est.meth.mort %>% mutate(
  var=case_when(
  method.mort>=1 & method.mort<=3 ~ "Based on VR data",
  method.mort>=4 & method.mort<=6 ~ "Using IHME estimates",
  method.mort>=7 & method.mort<=9 ~ "Based on CFR",
  method.mort==99 ~ "Country-specific customization"
  )
  ) %>%
  mutate(
    var = factor(var, levels = c("Based on VR data", "Based on CFR", "Using IHME estimates", "Country-specific customization"))
    )

cm.lst=unique(est.meth.mort$iso3[est.meth.mort$method.mort==2 | est.meth.mort$method.mort==5 |
                                est.meth.mort$method.mort==8])

rm.lst=unique(est.meth.mort$iso3[est.meth.mort$method.mort==3 | est.meth.mort$method.mort==6 |
                                est.meth.mort$method.mort==9])

method1=est.meth.mort[,.(iso3,var)]

f1.2.16_plot <- whomapper(
  X = method1, water_col = 'white',
  legend_title = 'Main method',
  #colours = brewer.pal(4, "Paired"),
  colours = brewer.pal(3, "Set3"),
  #colours= c("#2c7fb8","#bdd7e7","#88be3b","#c51b8a","#fbb4b9"),
  legend_pos = c(0.14,0.48)
  )


f1.2.16_plot_final <- f1.2.16_plot +
  add_marker(iso3=cm.lst, col="#f4a81d", size=2, shape=16) + 
  add_marker(iso3=rm.lst, col="#754d06", size=2, shape=16) + 
  guides(shape=FALSE) 
  # add_annotate("text", label="Country-specific dynamic model", x=-110, y=-29, size=I(2), include_coord = TRUE) +
  # add_annotate("text", label="Region-specific dynamic model", x=-112.5, y=-33, size=I(2)) +
  # add_annotate("point",  x=220, y=-29, color="#f4a81d", size=I(2)) +
  # add_annotate("point", x=217.25, y=-33, color="#754d06", size=I(2))
  
output_ggplot(f1.2.16_plot_final, dta, show_static=TRUE, pdf_csv_folder, save_csv, save_pdf)

```
<div class="footnote"> IHME: Institute of Health Metrics and Evaluation, Seattle, USA.</div> 

<div class="footnote"> VR: Vital registration of causes of deaths.</div> 

</div>




Country-specific details are available in the [Global tuberculosis report app](https://www.who.int/teams/global-tuberculosis-programme/data#app) and [country profiles](https://worldhealthorg.shinyapps.io/tb_profiles/).

Note : All of the percentage reductions or increases referred to in the text of this webpage were calculated based on unrounded numbers, as opposed to the rounded numbers that appear in the interactive graphics.

`r anch("refs")`

<hr style="border:1px solid gray20">

**References**


1. International statistical classification of diseases and health related problems (The) ICD-10. Geneva: World Health Organization; 2016 (https://icd.who.int/browse10/2016/en). 

2. Coronavirus (COVID-19) dashboard. Geneva: World Health Organization. (https://data.who.int/dashboards/covid19/cases).

3. AIDS info [website]. 2024 (http://aidsinfo.unaids.org/). 

4. Guidelines for Accurate and Transparent Health Estimates Reporting (GATHER) [website]. Geneva: World Health Organization; 2021 (http://gather-statement.org/). 

5. Global tuberculosis report 2023. Geneva: World Health Organization; 2023 (https://iris.who.int/handle/10665/373828). License: CC BY-NC-SA 3.0 IGO.

6. GBD results tool [website]. Global Health Data Exchange; 2020 (http://ghdx.healthdata.org/gbd-results-tool). 


<hr style="border:1px solid gray20">

**General disclaimers**
<br>
The designations employed and the presentation of the material in this publication do not imply the expression of any opinion whatsoever on the part of WHO concerning the legal status of any country, territory, city or area or of its authorities, or concerning the delimitation of its frontiers or boundaries. Dotted and dashed lines on maps represent approximate border lines for which there may not yet be full agreement.



```{r js_functions}
# Insert javascript file containing common Kendo number formatting functions ----
cat(writeLines(readLines(here("report/resources/gtbr_js.htm"))))
```


```{r, json_2_1} 
# renaming variable for Kendo UI
names(global) <- gsub('[.]', '_', names(global))

names(regional) <- gsub('[.]', '_', names(regional))

names(hest) <- gsub('[.]', '_', names(hest))

#names(excessmort) <- gsub('[.]', '_', names(excessmort))

names(income) <- gsub('[.]', '_', names(income))


# fig 1.2.4
tbhiv <- global$mort_h_num[global$year == 2019] / 1e3
top10$tbhiv <- c(rep(0, 12), tbhiv, rep(0, 7))
top10$n <- top10$deaths / 1000 + top10$tbhiv / 1000

# fig 1.2.5
names(unaids) <- gsub('[.]', '_', names(unaids))
hivd <- unaids[, .(sum(mort_hiv_num, na.rm = T)), by = year]
cod <- data.table(
  cause = c('HIV/AIDS', 'TB'),
  n = c((last(hivd$V1) - last(global$mort_h_num)) , last(global$mort_nh_num) ),
  tbhiv = rep(last(global$mort_h_num) , 2)
)
cod[, total := n + tbhiv]

# fig 1.2.6
m <- 1e5
ghiv.mort <-
  est[, addXY(mort.hiv / m, r.sd = mort.hiv.sd / m, weights = as.numeric(pop)), by =
        year]
dta_1_2_6 <- merge(global, ghiv.mort, by = 'year')
names(dta_1_2_6) <- gsub('[.]', '_', names(dta_1_2_6))

# fig 1.2.12
dta_1_2_12 <- est[iso3 %in% c('KHM','RUS','ZWE')]
dta_1_2_12[, mort.milestone := mort.num[16] * mort.milestone.vs.2015, by = iso3]
names(dta_1_2_12) <- gsub('[.]', '_', names(dta_1_2_12))


desired_age_order <- c("0-4", "5-14", "15-24", "25-34", "35-44", "45-54", "55-64", "65plus")
Mglobsplt$age <- factor(Mglobsplt$age, levels = desired_age_order, ordered = TRUE)
Mregsplt$age <- factor(Mregsplt$age, levels = desired_age_order, ordered = TRUE)
MPglobsplt$age <- factor(Mglobsplt$age, levels = desired_age_order, ordered = TRUE)
MPregsplt$age <- factor(Mregsplt$age, levels = desired_age_order, ordered = TRUE)


```


<script type="text/javascript">

/* JSON data objects for the figures */

var fig_1_2_1a_data = `r subset(global,year>=2010) %>% select(year,mort_num,mort_lo_num,mort_hi_num,mort_h_num,mort_h_hi_num,mort_h_lo_num,mort_nh_num,mort_nh_hi_num,mort_nh_lo_num, mort_milestone)  %>% toJSON("rows")`; 
var fig_1_2_1b_data = `r subset(global,year>=2010) %>% select(year,mort,mort_lo,mort_hi,mort_h,mort_h_hi,mort_h_lo,mort_nh,mort_nh_hi,mort_nh_lo)  %>% toJSON("rows")`; 


var fig_1_2_3_data = `r Mglobsplt %>%  select(sex,age,mort)  %>%  pivot_wider(names_from = sex, values_from = mort) %>% mutate(F = F*(-1)) %>% arrange(desc(age))%>% mutate(age=ifelse(age=="65plus","\u226565",as.character(age))) %>% toJSON("rows")`; 

var fig_1_2_3b_data = `r MPglobsplt %>%  select(sex,age,mort)  %>%  pivot_wider(names_from = sex, values_from = mort) %>% mutate(F = F*(-1)) %>% arrange(desc(age))%>% mutate(age=ifelse(age=="65plus","\u226565",as.character(age))) %>% toJSON("rows")`; 

var fig_1_2_3new_data = `r Mglobsplt.piv %>% arrange(age)%>% mutate(age=ifelse(age=="65plus","\u226565",as.character(age))) %>% toJSON("rows")`; 

var fig_1_2_3hnew_data = `r MPglobsplt.piv %>% arrange(age)%>% mutate(age=ifelse(age=="65plus","\u226565",as.character(age))) %>% toJSON("rows")`; 


var fig_1_2_4_data = `r top2021  %>% select(causename,deaths,tbhiv2) %>% toJSON("rows")`; 

var fig_1_2_5_data = `r cod %>% arrange(desc(n)) %>% toJSON("rows")`; 

var fig_1_2_6_data = `r subset(dta_1_2_6,year>=2010) %>% select(year,mort_h_num,mort_h_hi_num,mort_h_lo_num,mort_nh_num,mort_nh_hi_num,mort_nh_lo_num,r_num,r_lo_num,r_hi_num) %>% toJSON("rows")`; 

var fig_1_2_7_data = `r subset(regional,year>=2010) %>% select(region,year,mort_h,mort_h_hi,mort_h_lo,mort_nh,mort_nh_hi,mort_nh_lo)  %>% toJSON("rows")`; 

var fig_1_2_8_data = `r subset(regional,year>=2010) %>% select(entity = region,year,mort_num,mort_hi_num,mort_lo_num,mort_milestone)  %>% toJSON("rows")`; 

var fig_1_2_8a_data = `r subset(income,year>=2010) %>% dplyr::select(group,year,mort_num,mort_hi_num,mort_lo_num,mort_milestone) %>% toJSON("rows")`; 


var fig_1_2_11_data = `r subset(hest,year>=2010) %>% select(entity = country,year,mort_num,mort_hi_num,mort_lo_num,mort_milestone)  %>% toJSON("rows")`; 

var fig_1_2_9_afr = `r Mregsplt %>% filter(name == "African Region") %>% select(name,sex,age,mort) %>% pivot_wider(names_from = sex, values_from = mort) %>% mutate(F = F*(-1)) %>% arrange(desc(age)) %>% mutate(age=ifelse(age=="65plus","\u226565",as.character(age))) %>% toJSON("rows")`; 

var fig_1_2_9_amr = `r Mregsplt %>% filter(name == "Region of the Americas") %>% select(name,sex,age,mort) %>% pivot_wider(names_from = sex, values_from = mort) %>% mutate(F = F*(-1)) %>% arrange(desc(age)) %>% mutate(age=ifelse(age=="65plus","\u226565",as.character(age))) %>% toJSON("rows")`; 

var fig_1_2_9_sea = `r Mregsplt %>% filter(name == "South-East Asia Region") %>% select(name,sex,age,mort) %>% pivot_wider(names_from = sex, values_from = mort) %>% mutate(F = F*(-1)) %>% arrange(desc(age)) %>% mutate(age=ifelse(age=="65plus","\u226565",as.character(age))) %>% toJSON("rows")`; 

var fig_1_2_9_eur = `r Mregsplt %>% filter(name == "European Region") %>% select(name,sex,age,mort) %>% pivot_wider(names_from = sex, values_from = mort) %>% mutate(F = F*(-1)) %>% arrange(desc(age)) %>% mutate(age=ifelse(age=="65plus","\u226565",as.character(age))) %>% toJSON("rows")`; 

var fig_1_2_9_emr = `r Mregsplt %>% filter(name == "Eastern Mediterranean Region") %>% select(name,sex,age,mort) %>% pivot_wider(names_from = sex, values_from = mort) %>% mutate(F = F*(-1)) %>% arrange(desc(age)) %>% mutate(age=ifelse(age=="65plus","\u226565",as.character(age))) %>% toJSON("rows")`; 

var fig_1_2_9_wpr = `r Mregsplt %>% filter(name == "Western Pacific Region") %>% select(name,sex,age,mort) %>% pivot_wider(names_from = sex, values_from = mort) %>% mutate(F = F*(-1)) %>% arrange(desc(age)) %>% mutate(age=ifelse(age=="65plus","\u226565",as.character(age))) %>% toJSON("rows")`; 


var fig_1_2_9new_data_afr = `r mdisag_afr %>% arrange(age)%>% mutate(age=ifelse(age=="65plus","\u226565",as.character(age))) %>% toJSON("rows")`; 

var fig_1_2_9new_data_amr = `r mdisag_amr %>% arrange(age)%>% mutate(age=ifelse(age=="65plus","\u226565",as.character(age))) %>% toJSON("rows")`; 

var fig_1_2_9new_data_emr = `r mdisag_emr %>% arrange(age)%>% mutate(age=ifelse(age=="65plus","\u226565",as.character(age))) %>% toJSON("rows")`; 

var fig_1_2_9new_data_eur = `r mdisag_eur %>% arrange(age)%>% mutate(age=ifelse(age=="65plus","\u226565",as.character(age))) %>% toJSON("rows")`; 

var fig_1_2_9new_data_sea = `r mdisag_sea %>% arrange(age)%>% mutate(age=ifelse(age=="65plus","\u226565",as.character(age))) %>% toJSON("rows")`; 

var fig_1_2_9new_data_wpr = `r mdisag_wpr %>% arrange(age)%>% mutate(age=ifelse(age=="65plus","\u226565",as.character(age))) %>% toJSON("rows")`; 



var fig_1_2_12_data = `r subset(dta_1_2_12,year>=2010) %>% select(entity = country,year,mort_num,mort_hi_num,mort_lo_num,mort_milestone)  %>% toJSON("rows")`; 

</script>


```{js, echo=FALSE}


function tb_format_thou_2(n) { 
  //return thousands formatted to to significant figures
  nt = n/1000; 
  return num_spacer(Number(nt.toPrecision(2))*1000)
}

/* Functions to create the figures */

function createFig_1_2_1a(fig_ID) {
 
		$(fig_ID).kendoChart({
			dataSource: fig_1_2_1a_data,
			chartArea: {
				height: 500
			},	
			legend: {
				position: "bottom"
			},
      series: [{
				type: "line",
        name: "Total number TB deaths",
				field: "mort_num",
				color: "black",
				tooltip: {
					visible: true,
					template: "Total number of TB deaths (#= category #): #= Number((value/1e6).toPrecision(3)) # million"
				}
			},{
				type: "rangeArea",
				fromField: "mort_lo_num",
				toField: "mort_hi_num",
				color: "black",
				tooltip: {
					visible: true,
  				template: "95% uncertainty interval (#= category #): #= Number((value.from/1e6).toPrecision(3)) #–#= Number((value.to/1e6).toPrecision(3)) #"
				}
			},{
				type: "line",
        name: "TB deaths in people with HIV",
				field: "mort_h_num",
				color: "lightskyblue",
				tooltip: {
					visible: true,
					template: "TB deaths in people with HIV (#= category #): #= Number((value/1e6).toPrecision(3)) # million"
				}
			},{
				type: "rangeArea",
				fromField: "mort_h_lo_num",
				toField: "mort_h_hi_num",
				color: "lightskyblue",
				tooltip: {
					visible: true,
  				template: "95% uncertainty interval (#= category #): #= Number((value.from/1e6).toPrecision(3)) #–#= Number((value.to/1e6).toPrecision(3))  #"
				}
			},{
				type: "line",
        name: "TB deaths in people without HIV",
				field: "mort_nh_num",
				color: "blue",
				tooltip: {
					visible: true,
					template: "TB deaths in HIV-negative people (#= category #): #= Number((value/1e6).toPrecision(3)) # million"
				}
			},{
				type: "rangeArea",
				fromField: "mort_nh_lo_num",
				toField: "mort_nh_hi_num",
				color: "blue",
				tooltip: {
					visible: true,
  				template: "95% uncertainty interval (#= category #): #= Number((value.from/1e6).toPrecision(3)) #–#= Number((value.to/1e6).toPrecision(3)) #"
				}
			},{
        type: "line",
        dashType: "dash",
        field: "mort_milestone",
        color: "grey",
        markers: {
             size: 0,
             opacity: 0
           },
			tooltip: {
				visible: false
			}
      },],      
            render: function (e) {
            var draw = kendo.drawing;
            var padding = 100;
            var padding2 = 350;
            var element = this.element;
            var rect = new kendo.geometry.Rect([padding, padding2], [element.width() - 2 * padding, element.height() - 2 * padding2]);

            var text = new draw.Text("2025 milestone", [0, 0], { font: "bold 9px Verdana,Arial,sans-serif" });
            draw.align([text], rect, "start");
            draw.vAlign([text], rect, "start");
            e.sender.surface.draw(text);
          },
			valueAxis: {
        labels: {
					template: "#= axis_spacer(value/1e6) #"
				},
				title: {
					text: "Millions per year",
					visible: true
				},
        min: 0,
				line: {
					visible: false
				}
			},
			categoryAxis: {
			  max: 2022,
				field: "year",
				labels: {
					rotation: "auto",
          step: 2
				},
				majorGridLines: {
					visible: false
				},
				title: {
					text: "Year",
					visible: true
				}
			}

		});
}



function createFig_1_2_1b(fig_ID) {
 
		$(fig_ID).kendoChart({
			dataSource: fig_1_2_1b_data,
			chartArea: {
				height: 500
			},	
			legend: {
				position: "bottom"
			},
      series: [{
				type: "line",
        name: "Total TB deaths",
				field: "mort",
				color: "black",
				tooltip: {
					visible: true,
					template: "Total TB deaths (#= category #): #= Number((value).toPrecision(3)) # per 100&nbsp;000 population"
				}
			},{
				type: "rangeArea",
				fromField: "mort_lo",
				toField: "mort_hi",
				color: "black",
				tooltip: {
					visible: true,
  				template: "95% uncertainty interval (#= category #): #= Number((value.from).toPrecision(3)) #–#= Number((value.to).toPrecision(3)) #"
				}
			},{
				type: "line",
        name: "TB deaths in people with HIV",
				field: "mort_h",
				color: "lightskyblue",
				tooltip: {
					visible: true,
					template: "TB deaths in people with HIV (#= category #): #= Number((value).toPrecision(2)) # per 100&nbsp;000 population"
				}
			},{
				type: "rangeArea",
				fromField: "mort_h_lo",
				toField: "mort_h_hi",
				color: "lightskyblue",
				tooltip: {
					visible: true,
				  template: "95% uncertainty interval (#= category #): #= Number((value.from).toPrecision(2))  #–#= Number((value.to).toPrecision(2))  #"
				}
			},{
				type: "line",
        name: "TB deaths in people without HIV",
				field: "mort_nh",
				color: "blue",
				tooltip: {
					visible: true,
					template: "TB deaths in HIV-negative people (#= category #): #= Number((value).toPrecision(3)) # per 100&nbsp;000 population"
				}
			},{
				type: "rangeArea",
				fromField: "mort_nh_lo",
				toField: "mort_nh_hi",
				color: "blue",
				tooltip: {
					visible: true,
				  template: "95% uncertainty interval (#= category #): #= Number((value.from).toPrecision(3)) #–#= Number((value.to).toPrecision(3)) #"
				}
			},],      

			valueAxis: {
        labels: {
					template: "#= axis_spacer(value) #"
				},
				title: {
					text: "Rate per 100 000 population per year",
					visible: true
				},
        min: 0,
				line: {
					visible: false
				}
			},
			categoryAxis: {
				field: "year",
				labels: {
					rotation: "auto",
          step: 2
				},
				majorGridLines: {
					visible: false
				},
				title: {
					text: "Year",
					visible: true
				}
			}

		});
}





function createFig_1_2_3(fig_ID, data, category_field, min) {
  
		$(fig_ID).kendoChart({		
      dataSource: data,			
      chartArea: {
				height: 500
			},	  
			legend: {
				position: "bottom"
			},
			seriesDefaults: {
				type: "bar",
        stack: true,
        gap: 0.2
			},
			series: [{
				field: "M",
				color: "#F4A81D",
        tooltip: {
				visible: true,
				template: "Estimated number of deaths in males aged #= category # years: #= tb_format_thou_2(dataItem.M)#"
			}
			},{
				field: "F",
				color: "#6363C0",
        tooltip: {
				visible: true,
				template: "Estimated number of deaths in females aged #= category # years: #= tb_format_thou_2(dataItem.F*(-1))#"
			}
			},
              ],
			valueAxis: {
				labels: {
				template:  "#= kendo.toString(Math.abs(value),'n0').replace(/,/g, ' ') #",
				},
        title: {
					text: "Number of deaths"
				},
				line: {
					visible: false
				},
        axisCrossingValue: min,
        min: min
			},
			categoryAxis: {
				field: category_field,
				labels: {
					rotation: "auto"
				},
        title: {
					text: "Age group (years)"
				},
				majorGridLines: {
					visible: false
				}			}
		});
}


function createFig_1_2_4(fig_ID) {
		$(fig_ID).kendoChart({
			dataSource: fig_1_2_4_data,
			chartArea: {
				height: 600
			},	
			legend: {
				position: "bottom"
			},
			seriesDefaults: {
				type: "bar",
        stack: true
			},
			series: [{
				field: "deaths",
				color: "#00aaad",
        gap: 0.2,
        tooltip: {
				visible: true,
				template: "#= category #: #= Number((value/1e6).toPrecision(2)) # million"
			}
			},{
				field: "tbhiv2",
				color: "grey",
        tooltip: {
				visible: true,
				template: "#= category #, TB-HIV: #= Number((value/1e6).toPrecision(2)) # million"
			}
			},],
			valueAxis: {
				labels: {
					template: "#= axis_spacer(value/1e6) #"
				},
				title: {
					text: "Number of deaths (millions)"
				},
				line: {
					visible: false
				}
			},
			categoryAxis: {
				field: "causename",
				labels: {
					rotation: "auto"
				},
				majorGridLines: {
					visible: false
				}
			},

		});
}



function createFig_1_2_5(fig_ID) {
		$(fig_ID).kendoChart({
			dataSource: fig_1_2_5_data,
			chartArea: {
				height: 300
			},	
			legend: {
				position: "bottom"
			},
			seriesDefaults: {
				type: "bar",
        stack: true
			},
			series: [{
				field: "n",
				color: "#00aaad",
        gap: 0.2,
        tooltip: {
				visible: true,
				template: "#= category #: #= Number((value/1e6).toPrecision(2)) # million"
			}
			},{
				field: "tbhiv",
				color: "grey",
        tooltip: {
				visible: true,
				template: "TB-HIV: #= Number((value/1e6).toPrecision(2)) # million"
			}
			},],
			valueAxis: {
				labels: {
					template: "#= num_spacer(value/1e6) #"
				},
				title: {
					text: "Number of deaths (millions)"
				},
				line: {
					visible: false
				}
			},
			categoryAxis: {
				field: "cause",
				labels: {
					rotation: "auto"
				},
				majorGridLines: {
					visible: false
				}
			},

		});
}

function createFig_1_2_6(fig_ID) {
 
		$(fig_ID).kendoChart({
			dataSource: fig_1_2_6_data,
			chartArea: {
				height: 500
			},	
			legend: {
				position: "bottom"
			},
      series: [{
				type: "line",
        name: "HIV deaths",
				field: "r_num",
				color: "black",
				tooltip: {
					visible: true,
					template: "HIV deaths (#= category #): #= Number((value/1e6).toPrecision(2)) # million"
				}
			},{
				type: "rangeArea",
				fromField: "r_lo_num",
				toField: "r_hi_num",
				color: "black",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= Number((value.from/1e6).toPrecision(2))#–#= Number((value.to/1e6).toPrecision(2)) #"
				}
			},{
				type: "line",
        name: "TB deaths in people with HIV",
				field: "mort_h_num",
				color: "lightskyblue",
				tooltip: {
					visible: true,
					template: "TB deaths in people with HIV (#= category #): #= Number((value/1e6).toPrecision(2)) # million"
				}
			},{
				type: "rangeArea",
				fromField: "mort_h_lo_num",
				toField: "mort_h_hi_num",
				color: "lightskyblue",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= Number((value.from/1e6).toPrecision(2))  #–#= Number((value.to/1e6).toPrecision(2))  #"
				}
			},{
				type: "line",
        name: "TB deaths in people without HIV",
				field: "mort_nh_num",
				color: "blue",
				tooltip: {
					visible: true,
					template: "TB deaths in HIV-negative people (#= category #): #= Number((value/1e6).toPrecision(3)) # million"
				}
			},{
				type: "rangeArea",
				fromField: "mort_nh_lo_num",
				toField: "mort_nh_hi_num",
				color: "blue",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= Number((value.from/1e6).toPrecision(3)) #–#= Number((value.to/1e6).toPrecision(3)) #"
				}
			},],      

			valueAxis: {
        labels: {
					template: "#= axis_spacer(value/1e6) #"
				},
				title: {
					text: "Millions of deaths per year",
					visible: true
				},
        min: 0,
				line: {
					visible: false
				}
			},
			categoryAxis: {
				field: "year",
				labels: {
					rotation: "auto",
          step: 2
				},
				majorGridLines: {
					visible: false
				},
				title: {
					text: "Year",
					visible: true
				}
			}

		});
}


function createFig_1_2_7(fig_ID, data, filter, min, unit) {
 
  	// Filter the dataset on the country variable
		dataJSON = data.filter( element => element.region == filter);
  
		$(fig_ID).kendoChart({		
      dataSource: dataJSON,			
			chartArea: {
				height: 400
			},	      
      title: {
				text: filter,
				color: "black",
				font: "bold 14px  Arial,Helvetica,sans-serif",
        align: "center"
			},	  

			legend: {
				position: "bottom"
			},
      series: [{
				type: "line",
				field: "mort_nh",
				color: "blue",
        markers: { size: 4},
				tooltip: {
					visible: true,
					template: "TB mortality rate among HIV-negative people (#= category #): #= Number((value).toPrecision(2)) # per 100&nbsp;000 population"
				}
			},{
				type: "rangeArea",
				fromField: "mort_nh_lo",
				toField: "mort_nh_hi",
				color: "blue",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= Number((value.from).toPrecision(2)) #–#= Number((value.to).toPrecision(2)) #"
				}
			},{
				type: "line",
				field: "mort_h",
				color: "lightskyblue", 
        markers: { size: 4},
				tooltip: {
					visible: true,
					template: "TB mortality rate among people with HIV (#= category #): #= Number((value).toPrecision(2)) # per 100&nbsp;000 population"
				}
			},{
				type: "rangeArea",
				fromField: "mort_h_lo",
				toField: "mort_h_hi",
				color: "lightskyblue",
				tooltip: {
					visible: true,
				format: "{0}",
				template: "95% uncertainty interval (#= category #): #= Number((value.from).toPrecision(2)) #–#= Number((value.to).toPrecision(2)) #"
				}
			},],      
      
			valueAxis: {
        labels: {
					template: "#= axis_spacer(value) #"
				},
				title: {
					text: "Mortality rate per 100 000 population\nper year",
					visible: true
				},
       min: 0,
        axisCrossingValue: 0.1,
				line: {
					visible: false
				}
			},      

			categoryAxis: {
				field: "year",
				labels: {
					rotation: "auto",
          step: 2
				},
				majorGridLines: {
					visible: false
				},

				title: {
					text: "Year",
					visible: true
				}
			}

		});
}
 
function createFig_1_2_81112(fig_ID, data, filter, fig_height, min, unit) {
 
  	// Filter the dataset on the entity variable
		dataJSON = data.filter( element => element.entity == filter);
  
		$(fig_ID).kendoChart({		
      dataSource: dataJSON,			
			chartArea: {
				height: fig_height
			},	      
      title: {
				text: filter,
				color: "black",
				font: "bold 14px  Arial,Helvetica,sans-serif",
        align: "center"
			},	  

			legend: {
				position: "bottom"
			},
      series: [{
				type: "line",
				field: "mort_num",
				color: "black",
        markers: { size: 4},
				tooltip: {
					visible: true,
					template: "Total TB deaths (#= category #): #= tb_format_thou_2(value) #"
				}
			},{
				type: "rangeArea",
				fromField: "mort_lo_num",
				toField: "mort_hi_num",
				color: "black",
				tooltip: {
					visible: true,
				  template: "95% uncertainty interval (#= category #): #= tb_format_thou_2(value.from) #–#= tb_format_thou_2(value.to) #"
				}
			},{
        type: "line",
        dashType: "dash",
        field: "mort_milestone",
        color: "grey",
        markers: {
             size: 0,
             opacity: 0
           },
			tooltip: {
				visible: false
			}
      },],      
      
			valueAxis: {
        labels: {
					template: "#= axis_spacer(value/1e3) #"
				},
				title: {
					text: "Thousands per year",
					visible: true
				},
       min: 0,
        axisCrossingValue: 0,
				line: {
					visible: false
				}
			},      

			categoryAxis: {
				field: "year",
				labels: {
					rotation: "auto",
          step: 2
				},
				majorGridLines: {
					visible: false
				},

				title: {
					text: "Year",
					visible: true
				}
			}

		});
}


function createFig_1_2_8a(fig_ID, data, filter, fig_height, min, unit) {
 
  	// Filter the dataset on the entity variable
		dataJSON = data.filter( element => element.group == filter);
  
		$(fig_ID).kendoChart({		
      dataSource: dataJSON,			
			chartArea: {
				height: fig_height
			},	      
      title: {
				text: filter,
				color: "black",
				font: "bold 14px  Arial,Helvetica,sans-serif",
        align: "center"
			},	  

			legend: {
				position: "bottom"
			},
      series: [{
				type: "line",
				field: "mort_num",
				color: "black",
        markers: { size: 4},
				tooltip: {
					visible: true,
					template: "Total TB deaths (#= category #): #= tb_format_thou_2(value) #"
				}
			},{
				type: "rangeArea",
				fromField: "mort_lo_num",
				toField: "mort_hi_num",
				color: "black",
				tooltip: {
					visible: true,
				  template: "95% uncertainty interval (#= category #): #= tb_format_thou_2(value.from) #–#= tb_format_thou_2(value.to) #"
				}
			},{
        type: "line",
        dashType: "dash",
        field: "mort_milestone",
        color: "grey",
        markers: {
             size: 0,
             opacity: 0
           },
			tooltip: {
				visible: false
			}
      },],      
      
			valueAxis: {
        labels: {
					template: "#= axis_spacer(value/1e3) #"
				},
				title: {
					text: "Thousands per year",
					visible: true
				},
       min: 0,
        axisCrossingValue: 0,
				line: {
					visible: false
				}
			},      

			categoryAxis: {
				field: "year",
				labels: {
					rotation: "auto",
          step: 2
				},
				majorGridLines: {
					visible: false
				},

				title: {
					text: "Year",
					visible: true
				}
			}

		});
}



function createFig_1_2_9(fig_ID, data, filter, min) {
  
  
		$(fig_ID).kendoChart({		
      dataSource: data,			
      chartArea: {
				height: 500
			},	
			title: {
				text: filter,
				color: "black",
				font: "bold 14px  Arial,Helvetica,sans-serif",
        align: "center"
			},
			legend: {
				position: "bottom"
			},
			seriesDefaults: {
				type: "bar",
        stack: true,
        gap: 0.2
			},
			series: [{
				field: "M",
				color: "#F4A81D",
        tooltip: {
				visible: true,
				template: "Estimated number of deaths in males aged #= category # years: #= tb_format_thou_2(dataItem.M)#"
			}
			},{
				field: "F",
				color: "#6363C0",
        tooltip: {
				visible: true,
				template: "Estimated number of deaths in females aged #= category # years: #= tb_format_thou_2(dataItem.F*(-1))#"
			}
			},
              ],
			valueAxis: {
				labels: {
				template:  "#= kendo.toString(Math.abs(value),'n0').replace(/,/g, ' ') #",
				step: 2
				},
        title: {
					text: "Number of deaths"
				},
				line: {
					visible: false
				},
        axisCrossingValue: min,
        min: min
			},
			categoryAxis: {
				field: "age",
				labels: {
					rotation: "auto"
				},
        title: {
					text: "Age group (years)"
				},
				majorGridLines: {
					visible: false
				}			}
		});
}



function createFig_1_2_3a(fig_ID, data,maxy,title) {
   
		$(fig_ID).kendoChart({
			dataSource: data,
			chartArea: {
				height: 400
			},	
			legend: {
				position: "bottom"
			},
			title: {
				text: title,
				color: "black",
				font: "bold 14px  Arial,Helvetica,sans-serif",
        align: "center"
			},
			
      seriesDefaults: {
          spacing: 0.2
        },
			
			series: [{
        type: "line",
				field: "mort_F",
        errorLowField: "lo_F",
        errorHighField: "hi_F",
        errorBars: {color: "#6363C0", line: { width: 3 }},
        opacity: 0,
				color: "#6363C0",
        markers: {
          visible: true,
          background: "#6363C0",
          size: 8
        },
        
			tooltip: {
				visible: true,
				template: "Estimated number of TB deaths in females aged #= category # years: #= tb_format_thou_2(value) #"
			}
			}
			,{
				type: "rangeArea",
				fromField: "lo_F",
				toField: "hi_F",
				opacity: 0,
        color: "#6363C0",
				tooltip: {
				visible: true,
				format: "{0}",
				template: "95% uncertainty interval (age group #= category #): #= tb_format_thou_2(value.from) #\u2013#= tb_format_thou_2(value.to) #"
				}
			}
			],
			
			valueAxis: {
				labels: {
				template:  "#= kendo.toString(Math.abs(value),'n0').replace(/,/g, ' ') #",
				},
				title: {
					text: "Number of deaths"
				},
				line: {
					visible: false
				},
        min: 0,
        max: maxy,
			},
			categoryAxis: {
				field: "age",
				labels: {
					rotation: "auto"
				},
				title: {
					text: "Age group (years)"
				},
				majorGridLines: {
					visible: true
				}
			}

		});
}



function createFig_1_2_3b(fig_ID, data,maxy,title) {
   
		$(fig_ID).kendoChart({
			dataSource: data,
			chartArea: {
				height: 400
			},	
			legend: {
				position: "bottom"
			},
			title: {
				text: title,
				color: "black",
				font: "bold 14px  Arial,Helvetica,sans-serif",
        align: "center"
			},
			
      seriesDefaults: {
          spacing: 0.2
        },
			
			series: [{
        type: "line",
				field: "mort_M",
        errorLowField: "lo_M",
        errorHighField: "hi_M",
        errorBars: {color: "#F4A81D", line: { width: 3 }},
        opacity: 0,
				color: "#F4A81D",
        markers: {
          visible: true,
          background: "#F4A81D",
          size: 8
        },
        
			tooltip: {
				visible: true,
				template: "Estimated number of TB deaths in males aged #= category # years: #= tb_format_thou_2(value) #"
			}
			}
			,{
				type: "rangeArea",
				fromField: "lo_M",
				toField: "hi_M",
				opacity: 0,
        color: "#F4A81D",
				tooltip: {
				visible: true,
				format: "{0}",
				template: "95% uncertainty interval (age group #= category #): #= tb_format_thou_2(value.from) #\u2013#= tb_format_thou_2(value.to) #"
				}
			}
			],
			
			valueAxis: {
				labels: {
				template:  "#= kendo.toString(Math.abs(value),'n0').replace(/,/g, ' ') #",
				},
				title: {
					text: "Number of deaths"
				},
				line: {
					visible: false
				},
        min: 0,
        max: maxy,
			},
			categoryAxis: {
				field: "age",
				labels: {
					rotation: "auto"
				},
				title: {
					text: "Age group (years)"
				},
				majorGridLines: {
					visible: true
				}
			}

		});
}



```


```{js, echo=FALSE}

$(document).ready(function () {
createFig_1_2_1a("#fig_1_2_1a");  
createFig_1_2_1b("#fig_1_2_1b");


createFig_1_2_3a("#fig_1_2_3a",fig_1_2_3new_data,140000)
createFig_1_2_3b("#fig_1_2_3b",fig_1_2_3new_data,140000)

createFig_1_2_3a("#fig_1_2_3ha",fig_1_2_3hnew_data,50000)
createFig_1_2_3b("#fig_1_2_3hb",fig_1_2_3hnew_data,50000)

createFig_1_2_4("#fig_1_2_4");
createFig_1_2_5("#fig_1_2_5");
createFig_1_2_6("#fig_1_2_6");  

createFig_1_2_7("#fig_1_2_7_EUR",fig_1_2_7_data,"European Region",0.1,10);
createFig_1_2_7("#fig_1_2_7_AFR",fig_1_2_7_data,"African Region",1,10);
createFig_1_2_7("#fig_1_2_7_EMR",fig_1_2_7_data,"Eastern Mediterranean Region",0.1,30);
createFig_1_2_7("#fig_1_2_7_AMR",fig_1_2_7_data,"Region of the Americas",0.1,4);
createFig_1_2_7("#fig_1_2_7_WPR",fig_1_2_7_data,"Western Pacific Region",0.1,20);
createFig_1_2_7("#fig_1_2_7_SEAR",fig_1_2_7_data,"South-East Asia Region",0.1,80); 

createFig_1_2_81112("#fig_1_2_8_EUR",fig_1_2_8_data,"European Region",400,19e3,1.4);
createFig_1_2_81112("#fig_1_2_8_AFR",fig_1_2_8_data,"African Region",400,300e3,1.4);
createFig_1_2_81112("#fig_1_2_8_EMR",fig_1_2_8_data,"Eastern Mediterranean Region",400,50e3,1.4);
createFig_1_2_81112("#fig_1_2_8_AMR",fig_1_2_8_data,"Region of the Americas",400,15e3,1.4);
createFig_1_2_81112("#fig_1_2_8_WPR",fig_1_2_8_data,"Western Pacific Region",400,60e3,1.4);
createFig_1_2_81112("#fig_1_2_8_SEAR",fig_1_2_8_data,"South-East Asia Region",400,400e3,1.4);


createFig_1_2_9("#fig_1_2_9_EUR",fig_1_2_9_eur,"European Region",-2000);
createFig_1_2_9("#fig_1_2_9_AFR",fig_1_2_9_afr,"African Region",-30000);
createFig_1_2_9("#fig_1_2_9_EMR",fig_1_2_9_emr,"Eastern Mediterranean Region",-10000);
createFig_1_2_9("#fig_1_2_9_AMR",fig_1_2_9_amr,"Region of the Americas",-4000);
createFig_1_2_9("#fig_1_2_9_WPR",fig_1_2_9_wpr,"Western Pacific Region",-20000);
createFig_1_2_9("#fig_1_2_9_SEAR",fig_1_2_9_sea,"South-East Asia Region",-50000);

  createFig_1_2_3a("#fig_1_2_9a_AFR",fig_1_2_9new_data_afr,60000,"African Region")
  createFig_1_2_3b("#fig_1_2_9b_AFR",fig_1_2_9new_data_afr,60000,"African Region")
  
  createFig_1_2_3a("#fig_1_2_9a_AMR",fig_1_2_9new_data_amr,7000,"Region of the Americas")
  createFig_1_2_3b("#fig_1_2_9b_AMR",fig_1_2_9new_data_amr,7000,"Region of the Americas")

  createFig_1_2_3a("#fig_1_2_9a_EMR",fig_1_2_9new_data_emr,15000,"Eastern Mediterranean Region")
  createFig_1_2_3b("#fig_1_2_9b_EMR",fig_1_2_9new_data_emr,15000,"Eastern Mediterranean Region")

  createFig_1_2_3a("#fig_1_2_9a_EUR",fig_1_2_9new_data_eur,5000,"European Region")
  createFig_1_2_3b("#fig_1_2_9b_EUR",fig_1_2_9new_data_eur,5000,"European Region")

  createFig_1_2_3a("#fig_1_2_9a_SEA",fig_1_2_9new_data_sea,100000,"South-East Asia Region")
  createFig_1_2_3b("#fig_1_2_9b_SEA",fig_1_2_9new_data_sea,100000,"South-East Asia Region")

  createFig_1_2_3a("#fig_1_2_9a_WPR",fig_1_2_9new_data_wpr,35000,"Western Pacific Region")
  createFig_1_2_3b("#fig_1_2_9b_WPR",fig_1_2_9new_data_wpr,35000,"Western Pacific Region")


createFig_1_2_81112("#fig_1_2_11_AGO",fig_1_2_11_data,"Angola",250,5e3,1.8);
createFig_1_2_81112("#fig_1_2_11_COG",fig_1_2_11_data,"Congo",250,1e3,2);
createFig_1_2_81112("#fig_1_2_11_IND",fig_1_2_11_data,"India",250,250e3,2);
createFig_1_2_81112("#fig_1_2_11_MNG",fig_1_2_11_data,"Mongolia",250,0.2e3,1.5);
createFig_1_2_81112("#fig_1_2_11_PAK",fig_1_2_11_data,"Pakistan",250,30e3,1.4);
createFig_1_2_81112("#fig_1_2_11_THA",fig_1_2_11_data,"Thailand",250,7e3,2);

createFig_1_2_81112("#fig_1_2_11_BGD",fig_1_2_11_data,"Bangladesh",250,20e3,1.8);
createFig_1_2_81112("#fig_1_2_11_PRK",fig_1_2_11_data,"Democratic People's\nRepublic of Korea",250,5e3,2);
createFig_1_2_81112("#fig_1_2_11_IDN",fig_1_2_11_data,"Indonesia",250,60e3,1.4);
createFig_1_2_81112("#fig_1_2_11_MOZ",fig_1_2_11_data,"Mozambique",250,5e3,2);
createFig_1_2_81112("#fig_1_2_11_PNG",fig_1_2_11_data,"Papua New Guinea",250,2e3,2);
createFig_1_2_81112("#fig_1_2_11_UGA",fig_1_2_11_data,"Uganda",250,8e3,1.7);

createFig_1_2_81112("#fig_1_2_11_BRA",fig_1_2_11_data,"Brazil",250,4e3,1.6);
createFig_1_2_81112("#fig_1_2_11_COD",fig_1_2_11_data,"Democratic Republic\nof the Congo",250,25e3,1.6);
createFig_1_2_81112("#fig_1_2_11_KEN",fig_1_2_11_data,"Kenya",250,15e3,2);
createFig_1_2_81112("#fig_1_2_11_MMR",fig_1_2_11_data,"Myanmar",250,10e3,2);
createFig_1_2_81112("#fig_1_2_11_PHL",fig_1_2_11_data,"Philippines",250,15e3,1.6);
createFig_1_2_81112("#fig_1_2_11_TZA",fig_1_2_11_data,"United Republic of Tanzania",250,10e3,2);

createFig_1_2_81112("#fig_1_2_11_CAF",fig_1_2_11_data,"Central African Republic",250,2e3,2);
createFig_1_2_81112("#fig_1_2_11_ETH",fig_1_2_11_data,"Ethiopia",250,10e3,2);
createFig_1_2_81112("#fig_1_2_11_LSO",fig_1_2_11_data,"Lesotho",250,3e3,2);
createFig_1_2_81112("#fig_1_2_11_NAM",fig_1_2_11_data,"Namibia",250,1e3,2);
createFig_1_2_81112("#fig_1_2_11_SLE",fig_1_2_11_data,"Sierra Leone",250,2e3,2);
createFig_1_2_81112("#fig_1_2_11_VNM",fig_1_2_11_data,"Viet Nam",250,5e3,2);

createFig_1_2_81112("#fig_1_2_11_CHN",fig_1_2_11_data,"China",250,25e3,1.6);
createFig_1_2_81112("#fig_1_2_11_GAB",fig_1_2_11_data,"Gabon",250,0.5e3,2);
createFig_1_2_81112("#fig_1_2_11_LBR",fig_1_2_11_data,"Liberia",250,1e3,2);
createFig_1_2_81112("#fig_1_2_11_NGA",fig_1_2_11_data,"Nigeria",250,60e3,1.4);
createFig_1_2_81112("#fig_1_2_11_ZAF",fig_1_2_11_data,"South Africa",250,20e3,2);
createFig_1_2_81112("#fig_1_2_11_ZMB",fig_1_2_11_data,"Zambia",250,4e3,2);

createFig_1_2_81112("#fig_1_2_12_KHM",fig_1_2_12_data,"Cambodia",250,2e3,2);
createFig_1_2_81112("#fig_1_2_12_RUS",fig_1_2_12_data,"Russian Federation",250,5e3,2);
createFig_1_2_81112("#fig_1_2_12_ZWE",fig_1_2_12_data,"Zimbabwe",250,3e3,2);


createFig_1_2_8a("#fig_1_2_8a_HIC",fig_1_2_8a_data,"High-income countries",400,19e3,1.4);
createFig_1_2_8a("#fig_1_2_8a_LIC",fig_1_2_8a_data,"Low-income countries",400,19e3,1.4);
createFig_1_2_8a("#fig_1_2_8a_LMC",fig_1_2_8a_data,"Lower-middle-income countries",400,19e3,1.4);
createFig_1_2_8a("#fig_1_2_8a_UMC",fig_1_2_8a_data,"Upper-middle-income countries",400,19e3,1.4);


});  

```


